### 2016-05-23 paper-Ro_Man-2016.tex
update figures

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\usepackage[pdftex]{graphicx}
\usepackage{caption}
\usepackage{subcaption}
%\graphicspath{}
%\usepackage[utf8]{inputenc}

\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{io} =[text centered, draw=black]
\tikzstyle{process} = [text centered, text width=1cm, draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]
\tikzstyle{darrow} = [thick,<->,>=stealth] 
\tikzstyle{io2} =[ text centered, draw=black]
\tikzstyle{process2} = [text width=2cm, draw=black]

\usepackage{paralist}
\usepackage{amsfonts}
\usepackage{hyperref}

\usepackage{url}
\usepackage{booktabs}

\usepackage{tikz}
\usetikzlibrary{positioning, calc}
\usetikzlibrary{shapes.geometric, arrows}
% COLORS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{color}
\definecolor{lightgray}{rgb}{0.95, 0.95, 0.95}
\definecolor{darkgray}{rgb}{0.4, 0.4, 0.4}
%\definecolor{purple}{rgb}{0.65, 0.12, 0.82}
\definecolor{editorGray}{rgb}{0.95, 0.95, 0.95}
\definecolor{editorOcher}{rgb}{1, 0.5, 0} % #FF7F00 -> rgb(239, 169, 0)
\definecolor{editorGreen}{rgb}{0, 0.5, 0} % #007C00 -> rgb(0, 124, 0)
\definecolor{orange}{rgb}{1,0.45,0.13}		
\definecolor{olive}{rgb}{0.17,0.59,0.20}
\definecolor{brown}{rgb}{0.69,0.31,0.31}
\definecolor{purple}{rgb}{0.38,0.18,0.81}
\definecolor{lightblue}{rgb}{0.1,0.57,0.7}
\definecolor{lightred}{rgb}{1,0.4,0.5}

\usepackage[draft,nomargin,footnote]{fixme}

\graphicspath{{figs/}}

%\usepackage{xspace}

\title{\LARGE \bf
Child-Robot Spatial Arrangement in a Learning by Teaching Activity}

\author{Wafa Johal$^{1}$, Alexis Jacq$^{1}$$^{2}$, Ana Paiva$^{2}$ and Pierre Dillenbourg$^{1}$  % <-this % stops a space
\thanks{}% <-this % stops a space
\thanks{$^{1}$ CHILI Lab,
        École Fédérale Polytechnique Lausanne, Switzerland
        {\tt\small firstname.lastname@epfl.ch}}%
\thanks{$^{2}$Instituto Superior T\'{e}cnico, University of Lisbon, Portugal
        {\tt\small firstname.lastname@inesc-id.pt}}%
}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
In this paper, we present an experiment in the context of a child-robot interaction where we study the influence of the spatial arrangement of the interaction on the child's focus of attention and the perception of the robot's performance in the ``''Co-Writer learning by teaching'' activity.
In this activity, the child teaches a Nao robot how to write by hand.
Typically in a educational child robot interaction, face-to-face spatial arrangements are usually set.
We explored two spatial conditions from Kendon's F-formation, the side-by-side and the face-to-face formations in a within subject experiment.
We estimated the gaze behaviour of the child and the consistency of their grading of the robot with regard to the robot's progress in writing. 
Even-though the demonstrations provided by children were not different between the two conditions (i.e. the robot's learning didn't differ),  the results showed that in the side-by-side condition children tended to be more indulgent with the robot's mistakes and to give it better feedback.
These results will allow us to take into account the variability of the judgement of the children relatively to the robot's performances.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

Application of robots in learning tasks has been increasing and has shown very promising results these past ten years. 
R-learning, in reference to e-learning, aims to design robots that would be helpful for learning.
In our research context, we consider the problem of how robots could enhance the effectiveness of education.
The potential of robots for individual adaptation and their ability to handle repetitive tasks makes them fit well in the educational world.
In that vein, the Co-Writer projects aims to help children with difficulties in handwriting \cite{hood2015when}.
It is based on the idea of ``learning by teaching''.
Under this paradigm, by teaching a robot (the Nao, in our case) how to write, children learn and improve their handwriting at the same time.
The activity also plays with the protégé effect, which makes children more motivated in practising for the robot than for themselves.

Previous studies in the Co-Writer project helped to develop a system that generates handwriting for the robot based on demonstrations from the child\cite{hood2015when}.
Case studies presented in \cite{jacq2016building} showed that children were able to stay engaged in long term interactions with repeated sessions within the Co-Writer activity in real pedagogical/therapeutic contexts.
These works proved to have a positive effect on the extrinsic motivation of the children when practising their handwriting, thanks to the protégé effect.

Authors in \cite{Matsuzoe2012} had a similar approach to learning by teaching with their Care Receiving Robot, who was being taken care of and taught by a child using physical interaction. 
In this study, authors chose to investigate handwriting (or drawing shapes) as well, but with a more physical based approach.
In their experimental setting, the child would teach handwriting to the robot by placing himself behind the robot and moving it's hand.
This study varied from other works in educational human-robot interaction in a way that the child was not only facing the robot but would act as a care-taker rather than a teacher.

As robots are entering our living space, they must adapt to our social norms.
These norms vary from politeness to unspoken social rules (as for instance, the personal space of a person).
In home environments, robots will be expected to perform their functions in a manner that is clearly understandable and predictable by the humans around.
This requires adaptive personalisation of the robot to the individual needs of the humans, but also to the task being performed currently.
Some previous studies have shown that the spatial setting with a robot was also a way to convey non-verbal messages and it serves to influence the relationship with the user\cite{Takayama2009,kristoffersson2013measuring}.

In this study, we want to explore the effect of spatial arrangement on the child-robot interaction within the Co-Writer activity.

\section{RELATED WORKS}
Spatial configuration is part of non-verbal cues of interpersonal stance.
Spatial arrangement is also a social signal that tells about the relationship between people.
Hall \cite{hall} has proposed to divided the interpersonal space according to the distance between the individuals (intimate - the closer, personal social and public).

Kendon \cite{Kendon2010} called the F-Formation a spatial-orientational system that aims to explain how people arrange themselves in a group interaction.
%According to Kendon, the way people arrangement themselves during an interaction reflects their engagement and involvement with the other persons. 
Kendon proposed three main F-formations : face-to-face (or vis-a-vis), L shape and side-by-side. These formations are illustrated by Figure \ref{fig:fformations}.
\begin{figure}
	\centering
	\begin{subfigure}{0.2\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{./figures/fformationff.png}
		\caption{Face to face}
		\label{fig:ff}
	\end{subfigure}%
	\begin{subfigure}{0.2\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{./figures/fformationL.png}
		\caption{L-Shape}
		\label{fig:lshape}
	\end{subfigure}
	\begin{subfigure}{0.2\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{./figures/fformationss.png}
		\caption{Side by Side }
		\label{fig:sidebside}
	\end{subfigure}
	\caption{Three possible Kendon's F-formations around a rectangular table}
	\label{fig:fformations}
\end{figure}

According to \cite{knapp2013nonverbal}, spatial arrangement and seating position can determine a person's role in a group.
It could also predispose the person to either competitive (face to face) or cooperative (side by side) mental setting in a shared task.

In educational child-robot interaction, quite often, the robot is placed as a tutor or a peer helping the child, and uses social capabilities to enhance the learning \cite{Saerbeck2010}.
Kanda \cite{Kanda2004} reported an experiment involving children that learns from a robot as they would from a peer.
In this other experiment,
\cite{Leyzberg2014} showed in a face to face setting that the physical presence of the robot produced measurable learning gains.
\cite{Kennedy2015} confirmed these results but recommended caution when applying social behaviours in a tutoring context. 
%explain why caution

Spatial arrangements have been studied in previous research\cite{Kuzuoka,huttenrauch2006investigating,vazquez2014spatial} in HRI, often as a challenge of path planning, or as a metric but less frequently as a non-verbal cue of communication in a learning task.

We propose to study the impact of spatial arrangement on the engagement of children in a handwriting task.
We performed a study in a school with 12 children performing the same teaching task under two conditions.
The child was either facing the robot (similar to other tutoring experiments) or was sitting next to the robot (as a peer teaching setup).

%We also wanted to study the influence of being in a group or alone when interacting with the robot.
%At the difference from the group configuration, in the group configuration, other children were witnessing the interaction with the child.

\section{METHOD}
The experiment took place in a primary school in Switzerland where children are taught in English.
In this experiment, we targeted children aged 5 who start handwriting at that age, but do not typically master it yet.
The children interacted with the robot under two conditions of the F-Formation in a random order.

Our goal was to investigate the effect of spatial arrangement on the interaction.
We expected that children would give better feedback to the
when teaching it in a side-by-side configuration for several reasons.
The side-by-side arrangement corresponds to a peer teaching arrangement, unlike the face-to-face teaching arrangement, which is more frequent seen in a teacher-student relationship.
Also, in the side-by-side formation, the robot and the child have the same perspective of the shared tablet on which they write.
Perspective sharing is an ability that facilitates mutual understanding \cite{berlin2006perspective}.
Hence by having the robot and the child side-by-side, the mutual understanding would be eased.

\subsection{Participants and Apparatus}
12 subjects (six girls) from the same classroom (aged 5 to 6 y.o.) participated to the within subject study.
The two considered F-formations for this experiment are presented in the Figure~\ref{fig:test}: face-to-face~\ref{fig:sub1} and side-by-side~\ref{fig:sub2}.
Children were presented the two conditions sequentially with and interval of three days in a counterbalanced manner.

\begin{figure*}
\centering
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{./figures/f2f_photot.png}
  \caption{Face to face spatial formation}
  \label{fig:sub1}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{./figures/s2s_photo2t.jpg}
  \caption{Side by Side spatial formation}
  \label{fig:sub2}
\end{subfigure}
 \caption{Experimental set-ups showing the \emph{with-me-ness} targets in the rectangles: orange - the writing tablet,  blue - the robot's head, green - the tablet used to select a word to teach to the robot}
\label{fig:test}
\end{figure*}
Apart from this change in the spatial setting, the interaction was kept the same. 
The children were told they had to teach the robot how to write some words. 
We briefly presented the two tablet interfaces and the interaction started.
Any word from a list displayed on the selection tablet could be picked by the child.
As the robot would start to write this word, it was set to be a very bad writer at the beginning of the first session for each child.
The child could then give a feedback to the robot by pressing either a thumbs up or a thumbs down button.
The child would then use a pen and demonstrate how to write the word and the robot would then rewrite the word using the demonstration of the child.
The generated writing of the robot was computed to be halfway from its previous writing state and the new demonstration provided by the child.

Several hypothesis were made concerning the influence of the spatial arrangement on the interaction.
We expected that the gaze behaviour of the child would vary according to the spatial configuration.
More gazing at the robot's head would be eased in the face-to-face condition.
Our research question was to measure the degree to which this also influenced the way children behaved as a teacher (were they more or less severe with the robot facing them)
Because of this influence of facing the robot and being able to see its head, we also believed that the robot would be perceived as more competent in the face to face condition and hence collect more positive feedback from the child.
Indeed some previous studies have shown that social appearance \cite{goetz2003matching} has an influence on the perceived competences.

The engagement of the child in the task can also be influenced by the arrangement.
For that particular aspect, we will measure the number of repetitions of each word, as well as the with-me-ness, which is discussed in the following section.

\subsection{With-me-ness}
The with-me-ness introduced in HRI by \cite{lemaignan2016realtime}, helps to set specific targets during each state of the interaction and to determine whether the user is looking at one of these targets or not.
The algorithm is based on the \textit{d-lib library} that helps to estimate the head pose of the user using a video from a webcam device for instance.
From this head pose, the visual focus of attention (VFoA) is computed.
The inclusion of the targets in the VFoA allows then to compute the with-me-ness.
In a simple way, the with-me-ness value will increase if the child looks at the specific set of targets, or else it decreases.
This measure allows us to see if the child is engaged in the interaction and is looking at the tablet or the robot's head when he/she should.
This measure is actually very close to the notion of synchrony already studied in HRI \cite{delaherche2012interpersonal}, where bounding between individuals is reflected by their ability to synchronize in the task (look at the same time at the same targets).

For this experiment, the targets where: "the observer"(a teacher or a teacher assistant),"the experimenter",  "the selection tablet"(the tablet used to pick a word) and the "tablet".
The with-me-ness is initially set to 0.5 and takes values from 0 to 1.
According to the state in which the activity is, (robot is writing, child is writing,\dots) the with-me-ness will be increased if the child looks at a target that is in the set of targets that are optimum within this particular state of the activity.
We record these targets and the with-me-ness at a frequency of about $1Hz$.

The targets are defined according to their spatial relation with the camera used (placed on the table at the robot's feet).
The position of the targets was changed according to the F-formation condition, but the camera stayed at the same position
Figure \ref{fig:test} shows these targets for the two conditions.

%todo boucle avec Hypo

\subsection{Reward Mechanism}
The tablet interface on which the child and the robot practice their handwriting also contains two buttons that allow the child to give a positive (green thumb-up) or negative (red thumb-down) feedback to the robot's handwriting.
After every trial of the robot, the child could click on these feedback buttons as much as he/she wants.
We monitored each of these clicks.

These clicks aimed to assess the child's perception of the robot's progress.
When converging to a better writing we expect the child to give better feedback.
However, these buttons could also be used by the child as an encouragement method and children could give a positive feedback even though the robot didn't make progress.

\subsection{Performances}
As the child was managing which word the robot would learn, he/she could also provide as many demonstrations as he/she wanted.
The child was also free to change the word when satisfied with the robot's performance.

\subsubsection{Response Time and Writing Time}
We recorded the time spent on the writing and the response time for each word demonstrated by the children.
We also monitored the number of demonstrations provided by the child for each word.
These measure were cues to how dedicated the child was in the task

\subsubsection{Writing Score for the Robot}
%Regularity of writing in time
Since the learning algorithm took as input the child's demonstration, if the child provided repeatedly the exact same demonstration, then the robot would converge faster to their handwriting.
This score is hence a hint on children's regularity, with the underlying assumption that a child is sure of the fact that his handwriting is more regular.
Also, since children are teaching the robot, they might judge the robot's handwriting according to their provided demonstration.

The score for the generated letter by the robot was obtained by the distance between the same demonstrated letter by the child and the generated letter.
We finally computed the evolution of this score compare to the previous for each demonstrated word.
We then represented the evolution of the robot's handwriting with different states:
'S=': the first trial of this word by the robot, 'S-': the score decreases and 'S+' the score increases.

\section{RESULTS}
%The effect of IV on DV was statistically significant ($F_{a,b}= C, p<d*$). 
%The effect of IV on DV was not statistically significant ($F_{a,b}= C, ns$). 
In this section, we present results from this study on 12 children that interacted with the robot twice in the two conditions (side-by-side and face-to-face).
We studied the impact on the F-formation on the reward given with the buttons, the performance and the with-me-ness.

\subsection{Reward Mechanism}
Children well understood that they were teaching the robot and often gave bad scores for the first untrained trial of the robot.
Children gave feedback with an average of 3 feedbacks per demonstration.
Figure \ref{fig:uf_nb} shows the evolution of the sum of feedbacks given for all repeated word for each demonstration.
We noticed that the feedback is first negative and grows towards a positive feedback after each demonstration.
Children usually gave a positive feedback just before changing the word taught to the robot.

  \begin{figure}
  	\centering
  	\includegraphics[width=0.8\linewidth]{figures/plots/evoldemo}
  	\caption{Evolution of Feedback along the demonstration index}
  	\label{fig:uf_nb}
  \end{figure}

 The effect of the F-Formation on the feedback from the child was statistically significant ( $F_{1,272}= 4.396, p<0.05$). As, illustrated on Figure \ref{fig:userfeedback_fformation_se}, the average feedback per demonstration was greater($M=0.03, SD=1, N=147$) for the side-by-side condition compare to the face-to-face condition ($M=-0.23, SD=0.98, N=138$).
 \begin{figure}
 	\centering
 	\includegraphics[width=0.6\linewidth]{figures/plots/userfeedback_fformation_se}
 	\caption{Feedback According to the F-Formation}
 	\label{fig:userfeedback_fformation_se}
 \end{figure}

Children kept giving feedback along the interaction and no drop in the number of feedbacks was observed during the experiment.
They took their duty to teach the robot seriously in that way.

We also observe an order effect showing that children tended to give higher feedback in the second session compare to the first one.
%todo see if add significance here 
This can simply be explained by the fact that the robot's learning state was progressive between the two session. 
The robot didn't start to learn from scratch in the second session but had already some knowledge from the first session with this same child.

\subsection{With-me-ness}
Generally, children understood the dynamics of the interaction, as in general the with-me-ness stayed above 0.5 throughout the interaction (started et 0.5 but always finished above).
\begin{figure}	
\centering
\includegraphics[width=0.6\linewidth]{figures/plots/withmeness_fformation_ci}
\caption{With-me-ness According to the F-Formation (Means and Confidence Intervals)}
\label{fig:withmeness_fformation_ci}
\end{figure}
 The effect of the F-Formation on the with-me-ness of the child was statistically significant ( $F_{1,15983}= 293.2, p<0.001$). The with-me-ness was greater($M=0.79, SD=0.18, N=7722$) for the face-to-face condition compare to the side-by-side condition ($M=-0.72, SD=0.21, N=8267$) (see Figure \ref{fig:withmeness_fformation_ci}).
 This result was expected, as the robot was facing the robot, its face was more visible for the child.

Again, we observed an order effect with the withmeness increasing between the two session. 
As children were more confordable with the system, it is logic that they started to learn the dynamics of the interaction knwoing when to look at the selection tablet, the writing tablet and the robot. 
%todo add stats
%	session	N	new_w	sd	se	ci
%	1	1	8886	0.7248584	0.2175269	0.002307596	0.004523421
%	2	2	7103	0.7965673	0.1691839	0.002007419	0.003935140
%Error: Within
%Df Sum Sq Mean Sq F value Pr(>F)    
%session       1   12.8  12.797   339.8 <2e-16 ***
%Residuals 15985  602.0   0.038

\subsection{Performances}
%\begin{figure}
%\centering
%\includegraphics[width=0.4\linewidth]{figures/plots/score_alone_se}
%\caption{Learning speed According to group or alone modality}
%\label{fig:score_alone_ci}
%\end{figure}
\subsubsection{Response Time and Writing Time}
Figure \ref{fig:times} shows on the left the average number of demonstration for each word thought to the robot for the two spatial arrangement conditions. 
There was no significant difference between the conditions even-though the average number of demonstration given by the child in the side-by-side condition seems higher than the face-to-face.

The centrer graph of Figure \ref{fig:times} shows the average response time for the demonstration provided by the children for the two spatial arrangement conditions.
The response time is the delay between the time when the robot finishes to write and the time when the child touches the tablet.
There was no significant difference between the conditions.

The writing time (right graph on Figure \ref{fig:times}) corresponds to the delay between the time when the robot finishes its trial and the time the child finishes its new demonstration or changes word. This time also include the moment in which the child can give a feedback via the buttons.
No significant difference in the writing time was found.

\begin{figure*}[ht]
	\centering
	\includegraphics[width=0.25\linewidth]{figures/plots/repetition_fformation_ci}
	\includegraphics[width=0.25\linewidth]{figures/plots/responseTime_fformation_ci}
	\includegraphics[width=0.25\linewidth]{figures/plots/writingTime_fformation_ci2}
	\caption{Number of demonstration per word (left), response time (center), writing time (right) - (Means and Confidence Intervals)}
	\label{fig:times}
	\end{figure*}

These results show that children when not spending more time per word in one condition of the other. 
The spatial condition didn't influence the involvement of the child in the task. 
% The effect of the group variable on the score of the child was statistically significant ( $F_{1,279}= 6.928, p<0.01$). The score was greater($M=0.41, SD=0.11, N=141$) in  the alone condition compare to the group condition ($M= 0.37, SD=0.11, N=140$).
%These results confirm the fact that the robot is better suited here for individual interaction. 
\subsubsection{Writing Score for the Robot}
There was no significant difference of score in the F-formation condition tested (side-by-side:$M=0.80, SD=0.08, N=135$, face-to-face: $M=0.79, SD=0.08, N=191$).
This result means that the children were teaching as well in the face-to-face condition as in the side-by-side condition.
However, results showed significant differences in terms of feedback given to the robot regarding the score of the robot.
Table~\ref{proba} shows the probabilities of transition from after the score increases and after the score decreases.
When the score increases, we see that in the face-to-face condition, the probability of positive feedback is smaller than in the side-by side condition.
On the contrary, the probability of having a negative feedback after the score having decreased is higher in the face-to-face condition.
We can also notice that when the score decreases, the probability of having a positive feedback is twice higher in the side-by-side condition

\begin{table}[ht!]
\centering
\caption{\small Probability of feeback given the score per word in  the two conditions : face-to-face / \emph{side-by-side}}
\label{proba}
\begin{tabular}{l|p{2cm}|p{2cm}l@{}}
\toprule
Score event & Positive Feedback & Negative Feedback \\
\midrule
Score Increases & 	0.16 / \emph{0.27}	& 0.46 / \emph{0.47} \\
Score Decreases &	0.15 / \emph{0.32}	& 0.42 / \emph{0.28} \\
\bottomrule
\end{tabular}
\end{table}
\begin{figure*}
	\centering
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{./figures/faceface_score}
		\caption{Face to face spatial formation}
		\label{fig:f2fstate}
	\end{subfigure}%
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{./figures/sidebiside_score}
		\caption{Side by Side spatial formation}
		\label{fig:sbsstate}
	\end{subfigure}
	\caption{Transitions states from Score evolutions ($S+$ : score increases compared to previous score, $S-$: score decreases compare to the previous) to children's feedback}
	\label{fig:state}
\end{figure*}

All the transitions between the scores and the feedback buttons are illustrated on Figure \ref{fig:state}
Children were displaying a more positive attitude towards the robot when placed in side-by-side position.
This positive attitude was showed by rewarding more improvements of the robot and also penalising less the retrogression of the robot's writing.
Children even rewarded retrogression more often in the side-by-side condition.

\section{DISCUSSIONS AND CONCLUSIONS}
In this paper we presented a study that aimed to elicit the effect of spatial arrangement on a learning task between a robot and a child.
Spatial configuration has never to our knowledge been studied in terms of effect on the task but usually in the context of the social interaction.
In this experiment, we showed that this spatial arrangement could also have an influence in the task and on how the child was behaving as a teacher.
Our results showed that even if the robot's learning didn't vary between the side-by-side and the face-to-face condition, the feedback from the child was varying.
This suggests that the relationship between the child and the robot was different between the conditions.
These results are in line with the literature in spatial arrangement and non-verbal social cues.

We noticed that the children were expecting the robot to react to the feedback.
We are working on different form of reactions for future experiments to allow the robot to react according to the feedback from the child and also in perspective with the fairness of the feedback regarding the score of the robot.
For instance, if the child gives a positive feedback when the robot is actually improving, the robot should display a positive emotion.
If it is given a positive feedback but does not actually make any progress, he might express doubt to force the child to be more exigent.

Some limitations are however to be noticed. 
In this experiment, the robot was moving only when writing. 
Even if our results showed that the side-by-side elicited more positive feedback from the children, in the case where the robot would display some emotional reaction during the interaction, the same effect might not be observed. 
We plan to enhance the interaction with such reaction and to research the effect of such reactions on the child's perception of the robot's progresses. 
Long term studies might also bring more insights to theses results in order to see if the effect of the spatial arrangement sustain over time.
%discuss results regarding the literature, same ? different arguments for/against initial premise, limitations, etc.

%\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{ACKNOWLEDGMENT}
This research was partially supported by the Funda\c{c}\~{a}o para a Ci\^{e}ncia
e a Tecnologia (FCT) with reference UID/CEC/ 50021/2013, and by the Swiss
National Science Foundation through the National Centre of Competence in
Research Robotics.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrv}
\bibliography{fformation_cowriter}

\end{document}